Data Specialization for a Staged Lambda Calculus

### 2014-11-12 paper/semantics.tex
migrate overview section's text into semantics section

\section{Terminology and Overview}

@@

%We present an overview of the formal development along with our
%terminology, as summarized in \ref{fig:termSplitSummary}.

Our starting point is \lang, a typed, two-stage lambda calculus with products,
sums, and isorecursive types. This language is adapted from \cite{davies96}, and
is very similar to the staged languages considered in the partial evaluation and
metaprogramming literature. In this paper, we restrict our attention to a
two-stage language to simplify our presentation of the splitting algorithm, but
we expect that the techniques presented here extend to more stages.

The dynamic semantics of \lang\ differs from Davies's approach and
similar approaches in metaprogramming literature, because it avoids
duplication of staged work. Duplication takes place in meta
programming when a second-stage computation is silently duplicated by
the semantics.  In addition to leading to increases in run time, a
duplicating semantics makes it harder to reason about the splitting
algorithm, whose development and exposition is our main goal.  Apart
from the non-duplication, the dynamic semantics of \lang\ is
essentially identical to  Davies'.

We define the dynamics semantics of \lang\ not as an explicit
evaluation relation but as a relation that produces a term in an
unstaged, completely standard language, which we refer to as
\langmono\, because it is essentially lambda calculus extended with
product types. Given a \lang\ term, the dynamic semantics can be used
to reduce the term to a \langmono\ program, called {\em residual} by
using one application of stage-1 evaluation, one application of
reification, followed by one application of stage-2 evaluation.

\ur{Splitting...}

\crem{Fold previous section in here.}

@@

@@

>>>>

@@

@@

The input to our splitting algorithm is given in \lang, a typed, two-stage
lambda calculus with products, sums, and isorecursive types. This language is
adapted from \cite{davies96}, and is very similar to the staged languages
considered in the partial evaluation and metaprogramming literature. In this
paper, we restrict our attention to a two-stage language to simplify our
presentation of the splitting algorithm, but we expect that the techniques
presented here extend to more stages.

Our dynamic semantics of \lang\ differ from those given in \cite{davies96} and
the metaprogramming literature, because we avoid duplicating stage \bbtwo\
computations. In addition to being more efficient, our non-duplicating semantics
are easier to compare with our splitting algorithm, whose development and
exposition is our main goal.

%Apart from the non-duplication, the dynamic semantics of \lang\ is essentially
%identical to Davies'.

@@

\crem{fold in the next paragraph}

Given a \lang\ term, the dynamic semantics can be used
to reduce the term to a \langmono\ program, called {\em residual} by
using one application of stage-1 evaluation, one application of
reification, followed by one application of stage-2 evaluation.
